---
title: "Phoenix Sepsis EHR Implementation Notes"
subtitle: "__THIS IS A WORK IN PROGRESS__"
bibliography: ../references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The Phoenix criteria for pediatric sepsis and septic shock were developed using
an international data set of pediatric emergency department and hospital
encounters by predicting encounter mortality among those with suspected
infection using organ dysfunction-based data from the first 24 hours of the
encounter. Please see the original manuscripts for details.
[@sanchezpinto_2024_development;@schlapbach_2024_international].

The purpose of this document is to provide details about the expected inputs for
the scoring in order to assist EHR implementation. The phoenix R package and
Python module, along with the example SQL code assumes that the data have been
prepared as expected and returns scores accordingly.

<hr>

# Overview

Here is a quick checklist to summarize how the Phoenix package should be integrated within an EHR. This is to aid understanding and verification; users still need to review all guidance from this full document.

## Scoring Requirement

* Phoenix score is only valid in the case of suspected infection

## Exclusion Criteria

* Patients still in their original birth hospitalization
* Gestational age less than 37 weeks at encounter start
* Age greater than or equal to 18 years (216 months) at encounter start

## Data cleaning and preparation

* Verify all variables in expected unit; convert as necessary. E.g. age in months; lactate in mmol/L, etc.
* As appropriate replace invalid values with NULL; e.g. if age in months > 216 need to consider if should be scoring patient at all.
* No imputation for NULL/invalid values
* Last observation carry forward within appropriate window for each variable - max carry forward is from prior 24 hours.
* Variables not present or not known should be passed in as NULL

## Gotchas

* Phoenix package doesn't calculate mean arterial pressure (MAP); Calculate MAP from SBP and DBP if not directly available; if both cuff and arterial values present, arterial MAP should be used.
* Carefully review requirements of each input value. For example, only calculate SF ratio and pass into Phoenix package when SpO2 <= 97; only calculate and pass in PF ratio when FiO2 value is from the same time as or before the time of PaO2 lab.
* Use last known good value to when calling Phoenix functions.

<hr>

# General Data Formatting Notes

* `NULL` values are treated as non-score generating criteria. Do not populate null/unknown/not applicable values with estimates or outdated information.
* Values outside the ranges mentioned below are treated as `NULL`
* Before sending data to the package, please verify the data. Some values may provide warnings, others may generate errors. Verify that your code can handle warnings and errors.
* Ideally, both input values and time of the measurement are available so that last observation carried forward logic can be implemented with limits and conditions for age of inputs.

## Special Considerations for EHR Data

Phoenix was developed on a retrospective data set which provided the luxury of
easily looking backwards in the time within an encounter.  When applying the
Phoenix criteria to an active encounter there are several things to consider.

### Missing values
Missing values at a specific moment in time need to be dealt with in two steps:

1. Is there a prior known value?  If so, and the prior known value is within the
   lookback window, use the last known value.  That is,
   last-observation-carried-forward with a lookback window.

2. Consider the missing value to indicate a "healthy" value.  If using the R
   package or the Python module you need not do anything as the functions
   therein will return zero points in the score for the missing value.

Example: Say we know FiO~2~ at time T~0~ and we know lactate at time T~0~ as
well.

| time | FiO~2~ | Lactate |
|:--:  | :---   | :---    |
| T~0~ | 0.40   | 6.8     |

Later, at times T~2~ > T~1~ > T~0~ the lactate value is updated.  As such we
might see data like this:

| time | FiO~2~ | Lactate |
|:--:  | :---   | :---    |
| T~0~ | 0.40   | 6.8     |
| T~1~ |        | 7.3     |
| T~2~ |        | 7.1     |

It is important that when calculating the Phoenix score that the
missing value for FiO~2~ at time T~1~ is handled correctly:  The last known
value within the lookback window needs to be used.  That is: the data would look
like:

| time | FiO~2~                                                | Lactate |
| :--: | :---                                                  | :---    |
| T~0~ | 0.40                                                  | 6.8     |
| T~1~ | 0.40 (Assuming T~1~ - T~0~ < 6 hours; NULL Otherwise) | 7.3     |
| T~2~ | 0.40 (Assuming T~2~ - T~0~ < 6 hours; NULL Otherwise) | 7.1     |

If at time T~3~ FiO~2~ is updated:

 | time | FiO~2~                                                | Lactate                                              |
 | :--: | :---                                                  | :---                                                 |
 | T~0~ | 0.40                                                  | 6.8                                                  |
 | T~1~ | 0.40 (Assuming T~1~ - T~0~ < 6 hours; NULL Otherwise) | 7.3                                                  |
 | T~2~ | 0.40 (Assuming T~2~ - T~0~ < 6 hours; NULL Otherwise) | 7.1                                                  |
 | T~3~ | 0.80 (good new value)                                 | 7.1 (Assuming T~3~ - T~2~ < 6 hours; NULL Otherwise) |

<hr>

# Inclusion/Exclusion criteria for the Phoenix criteria development study

Inclusion Criteria

* Child < 18 years old
* Suspected infection

Exclusion Criteria

* Patients still in their original birth hospitalization
* Gestational age less than 37 weeks at encounter start
* Age greater than or equal to 18 years (216 months) at encounter start

# Suspected Infection

For a patient to have sepsis or septic shock based on the Phoenix criteria, they
must have a suspected infection.  A patient is considered to have a suspected
infection if they have:

1. a test for infection _ordered_, and
2. at least two doses of systemic antimicrobial medication(s).

Note that the test, e.g., a blood culture, need only be ordered.
Positive/negative test status is not needed.

Systemic antimicrobial medications are any antibiotic, antiviral, or antifungal
medication by any of the following routes:

* ASSUMED SYSTEMIC
* ENTERAL
* HEMODIALYSIS
* INTRA-ARTERIAL
* INTRACARDIAC
* INTRADUODENAL
* INTRAGASTRIC
* INTRAILEAL
* INTRAMUSCULAR
* INTRAVASCULAR
* INTRAVENOUS
* INTRAVENOUS BOLUS
* INTRAVENOUS DRIP
* NASOGASTRIC
* ORAL
* OROPHARYNGEAL
* PARENTERAL
* RECTAL
* SUBCUTANEOUS

Other non-systemic routes such as intraocular, nasal, otic, etc. should be excluded.

Example SQL code used for mapping medication routes when developing Phoenix can be found in
the project [GitHub archive](https://github.com/CU-DBMI-Peds/phoenix_sepsis_criteria/tree/main/harmonization/medication_mappings)
specifically the files:
[med_route_mapping.sql](https://github.com/CU-DBMI-Peds/phoenix_sepsis_criteria/blob/main/harmonization/medication_mappings/med_route_mapping.sql)
and
[clean_up_routes.sql](https://github.com/CU-DBMI-Peds/phoenix_sepsis_criteria/blob/main/harmonization/medication_mappings/clean_up_routes.sql).

<hr>

# General notes for all subscores

Reasonable values for observations and labs used during development of the Phoenix Sepsis Score are documented in
file [standard_names_and_units.csv](https://github.com/CU-DBMI-Peds/phoenix_sepsis_criteria/blob/main/harmonization/standard_names_and_units.csv)
within the project [GitHub archive](https://github.com/CU-DBMI-Peds/phoenix_sepsis_criteria).

"Look Back" for each input below indicates the length of time a last value should
be carried forward.

# Respiratory

Respiratory dysfunction scoring is based on PaO~2~/FiO~2~ (PFR), SpO~2~/FiO~2~
(SFR), and respiratory support.  While the scoring is based on SFR and PFR, it is
preferable to use separate SpO~2~, PaO~2~, and FiO~2~ values so data checks
can be performed before the score is determined.

## Inputs

### PaO~2~: Partial pressure of oxygen in arterial blood
* Units: mmHg
* Expected Values: [0, Inf)
* Caveats:
* Look Back: 6 hours

### SpO~2~: Peripheral oxygen saturation (pulse ox)
* Units: percentage
* Expected Values: [0, 100]
  * Expect integer values 0, 1, 2, ..., 100
* Caveats: values greater than 97 should _not_ be used for SFR
* Look back: 6 hours

### FiO~2~: Percent inspired oxygen
* Units: not applicable
* Expected Values: [0.21, 1.00]
  * expected floating point
  * 0.21 is room air
  * 1.00 is pure oxygen
* Caveats:
* Look back: 6 hours

### PFR: PaO~2~/FiO~2~
* Units: mmHg
* Expected Values [0, Inf)
  * Expect floating point
* Caveats:
  * PFR is only valid if the time of FiO~2~ value is the same as, or before,
    the time of PaO~2~ value.  For example, if FiO~2~ was reported at 11:31 and
    PaO~2~ was reported at 11:45 then the PFR would be valid to use.  If the
    FiO~2~ is then updated at 11:59, the PaO~2~ value from 11:45 would not be
    valid to use in the calculation of the PFR

    | time  | FiO~2~           | PaO~2~          | PFR    | Note:                                                        |
    |   --: | ---:             | ---:            | ---:   | ------:                                                      |
    | 11:31 | 0.30*            |                 |        |                                                              |
    | 11:45 | 0.30<sup>c</sup> | 118*            | 393.33 |                                                              |
    | 11:47 | 0.30<sup>c</sup> | 118<sup>c</sup> | 393.33 |                                                              |
    | 11:59 | 0.40*            | 118<sup>c</sup> |        | No PFR, FiO~2~ is newer than PaO~2~                        |
    | 12:10 | 0.40<sup>c</sup> | 152*            | 380    |                                                              |
    | 12:15 | 0.40<sup>c</sup> | 152<sup>c</sup> | 380    |                                                              |
    | 17:45 | 0.40<sup>c</sup> | 152<sup>c</sup> | 380    |                                                              |
    | 18:00 |                  | 152<sup>c</sup> |        | No PFR, FiO~2~ carry forward window expired                  |
    | 18:30 |                  |                 |        | No PFR, FiO~2~ and PaO~2~ carry forward windows have expired |

      : * New value <br>
      <sup>c</sup> value carried forward

* Look back: implied by the look back on PaO~2~ and FiO~2~ and the condition
  that FiO~2~ value is the same age or older than the PaO~2~ value.

### SFR: SpO~2~ / FiO~2~
* Units: none
* Expected Values: [0, 97/0.21]
  * Floating point
  * The upper limit includes the limitation on SpO~2~ ≤ 97
* Caveats:
  * SFR is only valid if the time of FiO~2~ value is the same as, or before,
    the time of SpO~2~ value.  For example, if FiO~2~ was reported at 11:31 and
    SpO~2~ was reported at 11:45 then the SFR would be valid to use.  If the
    FiO~2~ is then updated at 11:59, the SpO~2~ value from 11:45 would not be
    valid to use in the calculation of the SFR

     | time  | FiO~2~           | SpO~2~         | SFR   | Note:                                                        |
     | --:   | ---:             | ---:           | ---:  | ------:                                                      |
     | 10:15 | 0.25*            | 92*            | 368   |                                                              |
     | 10:20 | 0.25<sup>c</sup> | 98*            |       | No SFR, SpO~2~ > 97                                          |
     | 11:31 | 0.30*            | 98<sup>c</sup> |       | No SFR, SpO~2~ > 97 and FiO~2~ newer than SpO~2~           |
     | 11:45 | 0.30<sup>c</sup> | 87*            | 290   |                                                              |
     | 11:47 | 0.30<sup>c</sup> | 87<sup>c</sup> | 290   |                                                              |
     | 11:59 | 0.40*            | 87<sup>c</sup> |       | No SFR, FiO~2~ is newer than SpO~2~                        |
     | 12:10 | 0.40<sup>c</sup> | 85*            | 212.5 |                                                              |
     | 12:15 | 0.40<sup>c</sup> | 85<sup>c</sup> | 212.5 |                                                              |
     | 17:45 | 0.40<sup>c</sup> | 85<sup>c</sup> | 212.5 |                                                              |
     | 18:00 |                  | 85<sup>c</sup> |       | No SFR, FiO~2~ carry forward window expired                  |
     | 18:30 |                  |                |       | No SFR, FiO~2~ and SpO~2~ carry forward windows have expired |

      : * New value <br>
      <sup>c</sup> value carried forward

* Look back: implied by the look back on SpO~2~ and FiO~2~ and the condition
  that FiO~2~ value is the same age or older than the SpO~2~ value.

### ORS: Other Respiratory Support
* Units: not applicable
* Expected Values:
  * 0: No other respiratory support
  * 1: some form of respiratory support; may or may not include supplemental oxygen
* Caveats:
  * If FiO~2~ > 0.21 then that indicates oxygen support and thus ORS
* Look Back: 6 hours
  * If it has been more than six hours since last known data point indicating
    the patient has respiratory support then consider the patient to no longer
    have respiratory support.

### IMV: Invasive Mechanical ventilation
* Units: not applicable
* Expected Values:
  * 0: patient does not currently have invasive mechanical ventilation
  * 1: patient currently has invasive mechanical ventilation
* Caveats:
  * IMV due to surgery should not be used to assess organ dysfunction
* Look back: 6 hours
  * If it has been more than 6 hours since the last known data point
    indicating the patient has IMV, then consider the patient to no longer
    have IMV


## Scoring
Scoring is based on the lowest available PaO~2~/FiO~2~ (PFR) and
SpO~2~/FiO~2~ (SFR) with consideration for IMV and ORS.

```{r}
# R code: booleans are implicitly coerced to integers
cat(tail(as.character(body(phoenix::phoenix_respiratory)), 1), sep = "\n")
```

<hr>

# Cardiovascular

Cardiovascular dysfunction is based on three types of data:

1. Systemic Vasoactive medications
2. Lactate
3. Age adjusted mean arterial pressure

## Inputs

### Systemic Vasoactive Medications

There are six systemic vasoactive medications to consider:

1. dobutamine
2. dopamine
3. epinephrine
4. norepinephrine
5. milrinone
6. vasopressin

Vasoactive medications (dobutamine, dopamine, etc.) should only count if they
are continuous and systemic. Ideally, one would identify flowsheet or medication rows where the
dose (rate) is documented at least every hour for each infusion. In our dataset, epinephrine had
some non-systemic medication administrations that needed to be excluded.
Here are some *non-systemic* examples to exclude:

* epinephrine 0.3 mg/0.3ml inj soln prefilled syringe
* epinephrine hcl (nasal) 0.1 % nasal solution
* epinephrine hcl 2.25 % inh neb soln

Example SQL code used for mapping administration routes when developing Phoenix can be found in
the project [GitHub archive](https://github.com/CU-DBMI-Peds/phoenix_sepsis_criteria/tree/main/harmonization/medication_mappings)
specifically the files:
[med_route_mapping.sql](https://github.com/CU-DBMI-Peds/phoenix_sepsis_criteria/blob/main/harmonization/medication_mappings/med_route_mapping.sql)
and
[clean_up_routes.sql](https://github.com/CU-DBMI-Peds/phoenix_sepsis_criteria/blob/main/harmonization/medication_mappings/clean_up_routes.sql).

The overall metadata for vasoactives:

* Units: none
* Expected Values: integer values 0, 1, 2, 3, 4, 5, 6
  * (maximum vasoactive points is reached with 2)
* Caveats: see above
* Look back: 12 hours

### Lactate
* Units: mmol/L
* Expected Values: [0, 50]
* Caveats:
* Look back: 6 hours

### Age
* Units: Months
* Expected Values: [0, 216)
* Caveats:
* Look back:

### Mean Arterial Pressure (MAP)

There can be many sources of blood pressure.  We used the following hierarchy for
which value to use in the scoring.

1. Arterial MAP (invasive measurement)
2. Estimated MAP from arterial systolic (SBP) and arterial diastolic (DBP) pressures
3. Cuff MAP
4. Estimated MAP from cuff SBP and cuff DBP

MAP can be estimated as `(2/3) * DBP + (1/3) * SBP`

Expected metadata:

* MAP:
  * Units: mmHg
  * Expected Values: [1, 300]
  * Caveats:
    * If MAP is provided along with SBP and DBP then verify that DBP ≤ MAP ≤ SBP
  * Look back: 6 hours

* SBP:
  * Units: mmHg
  * Expected Values: [1, 300]
  * Caveats:
    * Verify that DBP ≤ SBP, we used a slightly more
      conservative check of (SBP - DBP) ≥ 1
  * Look back: 6 hours

* DBP:
  * Units: mmHg
  * Expected Values: [1, 200]
  * Caveats:
    * Verify that DBP ≤ SBP, we used a slightly more
      conservative check of (SBP - DBP) ≥ 1
  * Look back: 6 hours

**Step-By-Step Instructions:**
NOTE: these instructions are intended for prospective or real-time assessments
and differ from the implication used in the retrospective data for the
development of the Phoenix criteria.

1. Pick the score time. Example: ``We are scoring the patient at 10:15.''
2. Look back only as far as the blood-pressure look-back window allows. For
   published Phoenix, this is 6 hours.
3. Define 4 candidate estimators:
   a. Find the most recent arterial-line MAP within that window. If there is one, call it Candidate 1.
   b. Define Candidate 2.
      i. Find the most recent arterial-line SBP and arterial-line DBP within that window.
      ii. Only use arterial SBP and DBP to calculate MAP if they were recorded close enough together (within 10 minutes of each other). If they are too far apart in time, do not use them together.
      iii. If arterial SBP and DBP can be used together, calculate: MAP = one third SBP + two thirds DBP. Call this Candidate 2.
   c. Find the most recent cuff MAP within the window. If there is one, call it Candidate 3.
   d. Define Candidate 4.
      i. Find the most recent cuff SBP and cuff DBP within the window.
      ii. Only use cuff SBP and DBP to calculate MAP if they were recorded close enough together (within 10 minutes of each other). If they are too far apart in time, do not use them together.
      iii. If cuff SBP and DBP can be used together, calculate: MAP = one third SBP + two thirds DBP. Call this Candidate 4.

4. For each candidate, figure out how old it is at the score time.
   a. Ignore any candidate that does not exist.
   b. Choose the candidate that is newest.
      i. Candidate 1 and Candidate 3 have the same age as the reported value
      ii. Candidate 2 and 4: these are as old as the youngest component.  Example: SBP most recent reported value at 10:08; DBP last reported value at 10:03. Then the Candidate will be considered to have reported time of 10:08.

5. If two candidates are equally new (within one minute), use this tie-breaker order:
   a. Arterial-line MAP measured directly: Candidate 1.
   b. Arterial-line MAP calculated from arterial SBP and DBP: Candidate 2.
   c. Cuff MAP measured directly: Candidate 3.
   d. Cuff MAP calculated from cuff SBP and DBP: Candidate 4.

6. Use the Candidate selected in step 5 to assess the Phoenix score. If there
   are no usable candidates, then MAP is missing and gives 0 MAP points.

Simple Rule To Remember: Use the newest usable MAP. If there is a tie, prefer
arterial over cuff, and prefer directly measured MAP over calculated MAP.

Example: The candidate columns have the MAP value and the (age).

 | Time  | Arterial MAP | Cuff SBP | Cuff DBP | Candidate 1 | Candidate 4 | Selected Candidate | Note |
 | ----: | :---:        | :---:    | :---:    | :----       | :----       | :--:               | :--: |
 | 10:00 |              |          |          |             |             |                    |      |
 | 10:01 | 55           |          |          | 55 (0)      |             | 1                  | 1.   |
 | 10:02 | 56           |          |          | 56 (0)      |             | 1                  | 1.   |
 | 10:03 | 58           |          |          | 58 (0)      |             | 1                  | 1.   |
 | 10:04 |              | 63       | 47       | 58 (1)      | 52.3333 (0) | 1                  | 2.   |
 | 10:05 |              |          |          | 58 (2)      | 52.3333 (1) | 1                  | 2.   |
 | 10:06 |              |          |          | 58 (3)      | 52.3333 (2) | 1                  | 2.   |
 | 10:07 |              | 64       | 46       | 58 (4)      | 52 (0)      | 4                  | 3.   |
 | 10:08 | 57           |          |          | 57 (0)      | 52 (1)      | 1                  | 4.   |
 | 10:09 |              |          |          | 57 (1)      | 52 (2)      | 1                  | 4.   |
 | 10:10 |              |          |          | 57 (2)      | 52 (3)      | 1                  | 4.   |
 | 10:11 |              | 63       |          | 57 (3)      | 51.6667 (0) | 4                  | 5.   |
 | 10:12 |              |          |          | 57 (4)      | 51.6667 (1) | 4                  | 5.   |
 | 10:13 |              |          |          | 57 (5)      | 51.6667 (2) | 4                  | 5.   |
 | 10:14 |              |          |          | 57 (6)      | 51.6667 (3) | 4                  | 5.   |
 | 10:15 |              |          |          | 57 (7)      | 51.6667 (4) | 4                  | 5.   |
 | 10:16 |              |          |          | 57 (8)      | 51.6667 (5) | 4                  | 5.   |
 | 10:17 |              |          |          | 57 (9)      | 51.6667 (6) | 4                  | 5.   |
 | 10:18 |              |          |          | 57 (10)     | 51.6667 (7) | 4                  | 5.   |
 | 10:19 |              | 62       |          | 57 (11)     | ------      | 1                  | 6.   |

Notes:

1. The arterial MAP is the selected candidate because it is the only candidate and is aged zero minutes, well within in the look back window of six hours.
2. The selected candidate for use in the Phoenix scoring is Candidate 1, arterial MAP.  Although Candidate 4, the MAP approximated by cuff SBP and DBP, is younger, it is only younger by one minute. The age of measurement difference being only one minute, the two candidates are conceptually considered to be the same age and thus the Candidate 1, the arterial MAP is used when assessing Phoenix.
3. Candidate 4 is used here because the approximated MAP from cuff SBP and DBP is zero minutes old verse the four minute old arterial MAP.
4. Candidate 1 is used because it is the youngest and the preferable data source.
5. SBP was updated, DBP was not.  The approximate MAP based on cuff values can use the older DBP because it is within 10 minutes of the current SBP value.  The approximate MAP gets the same age as the youngest input, here the SBP.  This results in the cuff MAP being the youngest value by more than one minute and thus is the candidate used in the Phoenix assessment.
6. SBP was updated, DBP was not.  Because the SBP and DBP values are more than 10 minutes apart the estimated MAP is not to be used.  As such, Candidate 1, the arterial MAP is used based on data source preference and it is within the look back window.

## Scoring

Up to six points can be earned from cardiovascular dysfunction.  Up to two
points for vasoactives, up to two points for lactate, and up to two points for
MAP.

```{r}
# vas: number of vasoactive meds
# lct: lactate in mmol/L
# age: age in months
# map: mean arterial pressure in mmHg
cat(tail(as.character(body(phoenix::phoenix_cardiovascular)), 4), sep = "\n")
```

<hr>

# Coagulation

## Inputs

### Platelets
* Units: 1000/μL
* Expected Values: [0, Inf)
* Caveats:
* Look back: 24 hours

### INR
* Units: none
* Expected Values: [0, Inf)
* Caveats:
* Look back: 24 hours

### D-dimer
* Units: mg/L FEU
* Expected Values: [0, 500]
* Caveats:
* Look back: 24 hours

### Fibrinogen
* Units: mg/dL
* Expected Values: [0, Inf)
* Caveats:
* Look back: 24 hours

## Scoring

While there are four inputs and a patient could get a point for each one, the
final score is capped at two points.

```{r}
cat(tail(as.character(body(phoenix::phoenix_coagulation)), 2), sep = "\n")
```

<hr>

# Neurologic

## Inputs

### Glasgow Coma Scale (GCS)

Phoenix is based on the total GCS score.

* GCS (total):
  * Units: none
  * Expected Values: integer values 3, 4, 5, ..., 15
  * Caveats:
    * The total GCS is the sum of three component scores: eye, motor, and verbal.
    * We assumed that if we had an updated score for any one of the three components
      but not all, then the non-updated components were assumed to have not
      changed. In that case, we updated the time
      of the GCS total and all its components to the most recent update time for any
      of the components.
  * Look back: 12 hours

* GCS (eye):
  * Units: none
  * Expected Values: integer values 1, 2, 3, 4
  * Caveats:
  * Look back: 12 hours

* GCS (motor):
  * Units: none
  * Expected Values: integer values 1, 2, 3, 4, 5, 6
  * Caveats:
  * Look back: 12 hours

* GCS (verbal):
  * Units: none
  * Expected Values: integer values 1, 2, 3, 4, 5
  * Caveats: We did not use the T value for intubation - just use the recorded GCS-V, i.e. 1 for 1T
  * Look back: 12 hours

### Pupils
* Units: none
* Expected Values:
  * 1: bilaterally fixed pupils
  * 0: one or zero fixed pupils
* Caveats:
* Look back: 6 hours

## Scoring

* At most 2 points.
* 1 point for GCS ≤ 10
* 2 points for bilaterally fixed pupils

A patient with a GCS ≤ 10 and bilaterally fixed pupils will have a score of 2.

```{r}
# fpl: 1 if bilaterally fixed pupils; 0 otherwise
cat(tail(as.character(body(phoenix::phoenix_neurologic)), 1), sep = "\n")
```

<hr>

# Additional inputs to Phoenix-8 score

# Endocrine

## Inputs

### Glucose
* Units: mg/dL
* Expected Values: [5, 2000]
  * floating point
* Caveats:
* Look back: 12 hours

## Scoring
Scoring for Phoenix-8 only. 1 point if glucose < 50 or > 150

```{r}
cat(tail(as.character(body(phoenix::phoenix_endocrine)), 1), sep = "\n")
```

# Immunologic

## Inputs

### ANC
* Units: 1000 cells / mm³
* Expected Values: [0, Inf)
* Caveats:
  * Be careful with the units. The R package, Python module, and example SQL
    code expect units in 1000 cells / mm³.  Tables and other references might
    use cells / mm³
* Look back: 24 hours

### ALC
* Units: 1000 cells / mm³
* Expected Values: [0, Inf)
* Caveats:
  * Be careful with the units. The R package, Python module, and example SQL
    code expect units in 1000 cells / mm³.  Tables and other references might
    use cells / mm³
* Look back: 24 hours

## Scoring
At most one point

```{r}
cat(tail(as.character(body(phoenix::phoenix_immunologic)), 1), sep = "\n")
```

<hr>

# Renal

## Inputs

### Age (see notes in cardiovascular)

### Creatinine
* Units: mg/dL
* Expected Values: [0, 50]
* Caveats:
* Look back: 24 hours

## Scoring
Age-adjusted creatinine

```{r}
cat(tail(as.character(body(phoenix::phoenix_renal)), 1), sep = "\n")
```

<hr>

# Hepatic
## Inputs
### Total Bilirubin
* Units: mg/dL
* Expected Values: [0, 100]
  * floating point
* Caveats:
* Look back: 24 hours

### ALT
* Units: IU/L
* Expected Values: [0, Inf)
* Caveats:
* Look back: 24 hours

## Scoring
0 or 1 points

```{r}
cat(tail(as.character(body(phoenix::phoenix_hepatic)), 1), sep = "\n")
```

<hr>

# References
